---
output:
  html_document: default
  word_document: default
---
Course Project 1 Reproducible Research Tony Schollum
=======================================================

##Step 1 Read in the dataset

```{r}
activity<-read.csv("activity.csv")
```
set the environment
```{r,echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
```


Explore the data
```{r}
head(activity)
summary(activity)
```

It appears that we have to change the date column into a date format, the best option here is lubridate

```{r}
activity$date<-ymd(activity$date)
length(unique(activity$date))
```

and there are a number of missing steps counts to fix
```{r}
activity_na<-subset(activity, !is.na(activity$steps))
```

##Step 2 Histogram of Total Number of Steps taken per Day

```{r}
SPD<-tapply(activity_na$steps, activity_na$date, sum, na.rm=TRUE, simplify = T)

hist(x = SPD, col = rgb(0.8, 0.22,0,0.3), breaks = 30, xlab = "Steps", ylab = "Frequency", main = "Total Number of Steps Taken Per Day")
```

##Step 3 Mean and Median Steps per Day
Use of simple functions
```{r}
mean(SPD)
median(SPD)
```
## Step 4 Average Daily Activity Pattern
```{r}
Interval<- tapply(activity_na$steps, activity_na$interval, mean, na.rm = TRUE, simplify = T)
Int <- data.frame(interval = as.integer(names(Interval)), avg = Interval)
with(Int,
     plot(interval,
          avg,
          type = "b",
          col = "blue",
          xlab = "5 minute intervals",
          ylab = "avg steps in intervals"))
```

## Step 5, Interval showing the maximum number of steps```{r}
From the plot it appears to be around 800
```{r}
max_steps <- max(Int$avg)
Int[Int$avg ==max_steps,]
```

## Step 6, Code to impute missing data
```{r}
sum(is.na(activity$steps))
```

To calculate the missing data we can use the mean for the interval that is missing
```{r}
activity$CompleteSteps <- ifelse(is.na(activity$steps), round(activity$steps[match(activity$interval, activity$interval)],0), activity$steps)
```

```{r}
activityImp <- data.frame(steps=activity$CompleteSteps, interval=activity$interval, date=activity$date)
head(activityImp)
```
Step 7 use Weekday functionto see if there are differences in weekends to weekdays
```{r}
WW <- mutate(activity, weektype = ifelse(weekdays(activity$date) == "Saturday" | weekdays(activity$date) == "Sunday", "weekend", "weekday"))
WW$weektype <- as.factor(WW$weektype)
```
Calculate 5 min intervals and use ggplot 
```{r}
interval_full <- WW %>% group_by(interval, weektype) 
FM <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) + geom_line() + facet_wrap(~weektype, ncol = 1, nrow=2)
 print(FM)
```
